A scenario for a Planckian standard kilogram

### 2017-07-04 manuscript_gravitationalconstant.tex
revised introduction

\title{A thought-experiment for a Planckian standard kilogram}
% In a long title you can use \\ to force a line break at a certain location.

@@

\begin{abstract}
We present a new idea how the gravitational constant could be measured more precisely and if it could lead to a true Planckian basis for the SI unit system of measurement.
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

\maketitle % title page is now complete

@@

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For an simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[citation].

The proposed redefinition of the kilogram is as an explicit value for
$h$ and $c$, but we can think of it as
\begin{equation}
\mathrm{kg} \equiv 1.3563925\ldots \times 10^{50} \, \mathrm{s}^{-1}
                   \times \frac{h}{c^2}
\end{equation}
with the exact value to be decided.  As we can see, the kilogram is
coupled to the definition of a second.  That is not a problem in
practice, as the second is standardized orders of magnitude better
than any other unit.  Nonetheless, it is interesting to consider a
logical further step of decoupling the kilogram from other
definitions.  This could be done by introducing the gravitational
constant, since $\sqrt{hc/G}$ has dimensions of mass.  Historically,
there have been several proposed unit systems that included $G$ in
some way,\cite{Tomilin1999} but the idea of including both gravity and
quantum in a unit system comes from Planck\cite{Planck99}.  In the
process of working towards the radiation law, Planck realized that
$c,G,h$ and the Boltzmann constant $k_B$ (in his notation
$G\rightarrow f$, $h\rightarrow b$, $h/k_B\rightarrow a$) naturally
define units for mass, length, time, and temperature.  (Planck
remarked that such units would be meaningful even in non-human and
extraterrestrial cultures.)

\begin{equation}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}

In the late 19th century, Max Planck was working on reports about the irreversible process of radiation. He worked out a new entropy law to explain the empirical Wien law of radiation. This required him to introduce two new physical constants, which nowadays are known as the Planck constant and the Boltzmann constant . As a consequence, he described in 1899 a unit system which is based on fundamental natural constants only and thus free of anthropocentric arbitrariness. He pointed out the advantage of such a system in the following words:\cite{Planck99}\\

\begin{quote} ...ihre Bedeutung f\"{u}r alle Zeiten und f\"{u}r alle, auch au{\ss}erirdische und au{\ss}ermenschliche Kulturen notwendig behalten und welche daher als  ``nat\"{u}rliche Ma{\ss}einheiten" bezeichnet werden k\"{o}nnen... \end{quote}

\begin{quote} ...These necessarily retain their meaning for all times and for all civilizations, even extraterrestrial and non-human ones, and can therefore be designated as ``natural units"... \end{quote}

Planck units are defined in such a way that five universal physical constants are given by a numerical value of 1 when expressed in these units. Today the five constants are the speed of light in vacuum $c$, the gravitational constant $G$, the reduced Planck constant $\hbar$, the Boltzmann constant $k_B$ and the Coulomb constant $\left(4 \pi \epsilon_0\right)^{-1}$. Planck had defined them in a somewhat different form and did not adopt any electromagnetic unit. But the natural extension is to include the Coulomb constant, another possibility is to use the elementary charge $e$ to define a basic unit of charge \cite{Tomilin99}.\\
These constants can each be linked with a fundamental theory in physics that is widely accepted and universal: $\hbar$ with quantum mechanics, $c$ with special relativity, $G$ with general relativity, $k_B$ with thermodynamics and statistical mechanics, and $e$ or $\left(4 \pi \epsilon_0\right)^{-1}$ with electrodynamics.\\

The International System of Units (SI) is the commonly used system of measurement. Its beginnings go back to the French Revolution in 1799 as a precursor of the metric system was implemented. With the signing of the Convention of the Metre in 1875, new international prototypes of the metre and the kilogram were established. In the early twentieth century electromagnetic units joined the system and it underwent constant improvement, until in 1960 the modern form, called the Système International d'Unités (SI), was launched. Since the Mole joined the system of measurement in 1971, the SI unit system is in its present form.\\

To come away from the definitions based on artefacts, specific physical states and idealized experiments and because it is technologically possible to measure physical constants to high accuracy, the International Committee for Weights and Measures (CIPM) has proposed new definitions of the SI base units, which likely will be adopted in fall 2018 at the next General Conference on Weights and Measures. The revision of the SI unit system in 2018 goes towards a Planckian based system, since the values of seven physical constants get fixed and thereby define the seven base units. 
To get a truly Planckian, hence ``most natural", basis for the SI, one would have to exchange the ground state hyperfine splitting frequency of the caesium-133 atom $\Delta \nu_{\mathrm{Cs}}$ with the gravitational constant $G$ and then define the second by  
\begin{equation}
\mathrm{s} = \sqrt{\frac{h \, G}{c^5}}.
\end{equation}

Because these four natural constants would have fixed numerical values, this definition is exact. Such a definition would lead to an SI system that is based on fundamental natural constants and therefore stable over time and realizable everywhere. Also completely new and more precise measurements could be designed without having to revise the definitions of the units \cite{Fischer16}. The big problem is, that the uncertainty in the measurements of $G$ is very large large. In Fig.~\ref{fig:Gmeasurements} the latest and most precise measurements of the gravitational constant are plotted together with the recommended CODATA value and its uncertainty. At first glance one notices, that the measured values are not even in good agreement with each other.\\

The big problem that this approach faces, is that the current definition of the second which is based on the hyperfine splitting frequency in the Caesium 33 atom, is much more accurate. \\

>>>>

\title{A scenario for a Planckian standard kilogram}
% In a long title you can use \\ to force a line break at a certain location.

@@

\begin{abstract}
The kilogram could, in principle, be redefined as an integer multiple
of the Planck mass (or 4594xxxx$\times\sqrt{\hbar c/G}$) with no
significant change from its current value.  The practical hurdle is
measuring $G$ three or four orders of magnitude than presently known,
so as to find the unknown digits xxxx.  We study a possible scenario
for doing so with an artificial two-body orbit.  Requirements for
suppressing perturbing effects (notably tidal forces and radiation
pressure), and obtaining the desired signal-to-noise are estimated.
\end{abstract}

\maketitle

@@

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For an simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[2015AmJPh..83..913C].

The proposed redefinition of the kilogram is as an explicit value for
$h$ and $c$, but we can think of it as
\begin{equation}
\mathrm{kg} \equiv 1.3563925\ldots \times 10^{50} \, \mathrm{s}^{-1}
                   \times \frac{h}{c^2}
\end{equation}
with the exact value to be decided.  As we can see, the kilogram is
coupled to the definition of a second.  That is not a problem in
practice, as the second is standardized orders of magnitude better
than any other unit.  Nonetheless, it is interesting to consider a
logical further step of decoupling the kilogram from other
definitions.  This could be done by introducing the gravitational
constant, since $\sqrt{hc/G}$ has dimensions of mass.

Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum in a unit system comes from
Planck\cite{Planck99}.  In the process of working towards the
radiation law, Planck realized that $c,G,h$ and the Boltzmann constant
$k_B$ (in his notation $G\rightarrow f$, $h\rightarrow b$,
$h/k_B\rightarrow a$) naturally define units for mass, length, time,
and temperature.  (Planck remarked that such units would be meaningful
even in non-human and extraterrestrial cultures.)

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ times Planck's definitions.)  Extension to electrical
units is straightforward, but we will not discuss it here.  The Planck
length, time, and temperature in themselves are extreme scales, not
probed by any known physical process, and presumably the regime of
quantum gravity.  But extreme scales do not harm if they are only part
of some pression, as they are here.  The Planck mass is small but
microscopically measurable: $\sim20\mu\mathrm{g}$.  This is of the
same order as the uncertainty in the prototype kilogram itself.  Hence
the kilogram could be redefined as an integer multiple of the Planck
mass, without significantly perturbing its value.

The difficulty with Eq.~\eqref{eq:defs} is that all the expressions
involve $G$.  Given the current uncertainties in $G$,
Eq.~\eqref{eq:defs} would allow units to be standardized only to four
significant digits, which is obviously not enough.  Improvement by
several orders of magnitude would be needed, to make the Planckian
kilogram practicable.  In this paper, we study a scenario which may
make such an improvement possible.  A Planckian second not a prospect
with anything like present technology, because demands on the time
standard are more stringent still.

% In the late 19th century, Max Planck was working on reports about the irreversible process of radiation. He worked out a new entropy law to explain the empirical Wien law of radiation. This required him to introduce two new physical constants, which nowadays are known as the Planck constant and the Boltzmann constant . As a consequence, he described in 1899 a unit system which is based on fundamental natural constants only and thus free of anthropocentric arbitrariness. He pointed out the advantage of such a system in the following words:\cite{Planck99}\\

% \begin{quote} ...ihre Bedeutung f\"{u}r alle Zeiten und f\"{u}r alle, auch au{\ss}erirdische und au{\ss}ermenschliche Kulturen notwendig behalten und welche daher als  ``nat\"{u}rliche Ma{\ss}einheiten" bezeichnet werden k\"{o}nnen... \end{quote}

% \begin{quote} ...These necessarily retain their meaning for all times and for all civilizations, even extraterrestrial and non-human ones, and can therefore be designated as ``natural units"... \end{quote}

% Planck units are defined in such a way that five universal physical constants are given by a numerical value of 1 when expressed in these units. Today the five constants are the speed of light in vacuum $c$, the gravitational constant $G$, the reduced Planck constant $\hbar$, the Boltzmann constant $k_B$ and the Coulomb constant $\left(4 \pi \epsilon_0\right)^{-1}$. Planck had defined them in a somewhat different form and did not adopt any electromagnetic unit. But the natural extension is to include the Coulomb constant, another possibility is to use the elementary charge $e$ to define a basic unit of charge \cite{Tomilin99}.\\
% These constants can each be linked with a fundamental theory in physics that is widely accepted and universal: $\hbar$ with quantum mechanics, $c$ with special relativity, $G$ with general relativity, $k_B$ with thermodynamics and statistical mechanics, and $e$ or $\left(4 \pi \epsilon_0\right)^{-1}$ with electrodynamics.\\

% The International System of Units (SI) is the commonly used system of measurement. Its beginnings go back to the French Revolution in 1799 as a precursor of the metric system was implemented. With the signing of the Convention of the Metre in 1875, new international prototypes of the metre and the kilogram were established. In the early twentieth century electromagnetic units joined the system and it underwent constant improvement, until in 1960 the modern form, called the Système International d'Unités (SI), was launched. Since the Mole joined the system of measurement in 1971, the SI unit system is in its present form.\\

% To come away from the definitions based on artefacts, specific physical states and idealized experiments and because it is technologically possible to measure physical constants to high accuracy, the International Committee for Weights and Measures (CIPM) has proposed new definitions of the SI base units, which likely will be adopted in fall 2018 at the next General Conference on Weights and Measures. The revision of the SI unit system in 2018 goes towards a Planckian based system, since the values of seven physical constants get fixed and thereby define the seven base units. 
% To get a truly Planckian, hence ``most natural", basis for the SI, one would have to exchange the ground state hyperfine splitting frequency of the caesium-133 atom $\Delta \nu_{\mathrm{Cs}}$ with the gravitational constant $G$ and then define the second by  
% \begin{equation}
% \mathrm{s} = \sqrt{\frac{h \, G}{c^5}}.
% \end{equation}

% Because these four natural constants would have fixed numerical values, this definition is exact. Such a definition would lead to an SI system that is based on fundamental natural constants and therefore stable over time and realizable everywhere. Also completely new and more precise measurements could be designed without having to revise the definitions of the units \cite{Fischer16}. The big problem is, that the uncertainty in the measurements of $G$ is very large large. In Fig.~\ref{fig:Gmeasurements} the latest and most precise measurements of the gravitational constant are plotted together with the recommended CODATA value and its uncertainty. At first glance one notices, that the measured values are not even in good agreement with each other.\\

% The big problem that this approach faces, is that the current definition of the second which is based on the hyperfine splitting frequency in the Caesium 33 atom, is much more accurate. \\
